---
title: "assignment_8.4"
output:
  html_document:
    keep_md: true
  pdf_document: default
---

The objective of this analysis will be create a model capable of respond the
question *What is the quality of the exercise being realized*. To do so it will
utilize the
[Weight Lifting Exercises](https://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har)
dataset.

First, the caret library will be loaded and the data read. The *na.strings*
parameter is changed to account to digited values that doesn't represent a
real value. Also, the predicted class is transformed to a factor variable.

```{r}
library(caret)
```

```{r cache=TRUE}
data = read.csv('pml-training.csv', na.strings = c('', '#DIV/0!', 'NA'))
data$classe = as.factor(data$classe)
```

The first 7 columns representing time indicators and the user performing the
activity are removed as they aren't useful to predict new values.
Then, columns with a great number of null values are removed, as they can't be
properly estimated and used to make the predictions.

```{r}
data = data[, 8:160]
used_variables = which(colMeans(!is.na(data)) > .2)  # at least 20% without NA
dataremovenull = data[, used_variables]
```

After doing that, we can see that all predictor have appropriate types and that
all null values were removed.

```{r}
str(dataremovenull)
```
```{r}
colSums(is.na(dataremovenull))
```

After preparing the data, we will separate the data in training and test sets
using 80% of the data to train the model.

```{r}
set.seed(21)
in_train = createDataPartition(dataremovenull$classe, p = .8, list = F)
train_data = dataremovenull[in_train, ]
test_data = dataremovenull[-in_train, ]
```

To make the prediction, the random forest model were chosen.

```{r cache=TRUE}
set.seed(21)
model = train(classe ~ ., method = 'rf', data=train_data)
```

The model got a 100% accuracy on the training set and over 99% on the test set.
As such, the model got excellent results capable of making efficient predictions
of the quality of the exercise performed.

```{r}
confusionMatrix(train_data$classe, predict(model,train_data))
```

```{r}
confusionMatrix(test_data$classe, predict(model,test_data))
```
